---
title: "Figures used in manuscript and supplementary material"
author: "Lei Song"
date: "2022-10-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)

library(sf)
library(here)
library(raster)
library(terra)
library(ggplot2)
library(purrr)
library(pbmcapply)
library(dplyr)
library(tidyverse)
library(ggpubr)
library(stars)

data_path <- here("data")
rst_path <- file.path(data_path, "rasters")
vct_path <- file.path(data_path, "vectors")
hr_path <- file.path(data_path, "vars_home_range")
pat_path <- file.path(data_path, "vars_patch")
result_path <- here("results")
cnt_path <- file.path(data_path, "connectivity")
cnt_exp_path <- file.path(result_path, "connectivity")
fg_path <- here("docs/figures")
```

## Methodology
### Figure 1

Figure 1 is the figure of study area, datasets, etc.

```{r}
library(ggpattern)

# Read layers
study_area <- read_sf(
    file.path(vct_path, "mainland_tanzania.geojson"))
expert_range_map <- read_sf(
    file.path(data_path, "observations", "range_map.geojson")) %>% 
    filter(range == "Known") %>% 
    mutate(range = "Expert range map")
census_blocks <- read_sf(
    file.path(data_path, "observations", "census.geojson")) %>% 
    mutate(density_bin = mltools::bin_data(
        round(.$density, 2), 
        bins=8, binType = "quantile"))
zone_to_thin <- read_sf(
    file.path(data_path, "observations", "zone_to_thin.geojson")) %>% 
    mutate(zone = "Spatially-thin region")
occ <- read_sf(
    file.path(data_path, "observations", "ele_occurrence.geojson")) %>% 
    mutate(occ = "Occurrences")

# Plot
ggplot() +
    geom_sf(data = study_area, 
            fill = "transparent", color = "black") +
    geom_sf(data = zone_to_thin, 
            aes(fill = "Spatially-thin region"), 
            alpha = 0.5, show.legend = "polygon") +
    scale_fill_manual(
        NULL, values = "lightblue",
        breaks = 'Spatially-thin region',
        guide = guide_legend(
            override.aes = list(linetype = "blank", 
                                shape = NA))) +
    ggnewscale::new_scale_fill() +
    geom_sf_pattern(
        data = expert_range_map, 
        pattern = "magick",
        aes(pattern_type = 'Expert range map'),
        pattern_fill = 'black',
        pattern_scale = 2,
        pattern_aspect_ratio = 1,
        pattern_density = 0.1,
        color = "black", fill = "lightgrey") +
    scale_pattern_type_manual(
        NULL, values = "left45",
        breaks = "Expert range map",
        guide = guide_legend(
            override.aes = list(shape = NA))) +
    ggnewscale::new_scale_fill() +
    ggnewscale::new_scale_color() +
    geom_sf(data = census_blocks, 
            aes(fill = density_bin), 
            color = "black", size = 0.3) +
    scale_fill_brewer(
        "Population density\n(Census blocks)", 
        palette = "YlOrRd", direction = 1,
        guide = guide_legend(
            override.aes = list(linetype = "blank", 
                                shape = NA))) +
    ggnewscale::new_scale_fill() +
    ggnewscale::new_scale_color() +
    geom_sf(data = occ, aes(size = occ), 
            color = "white", fill = "blue",
            pch = 21, show.legend = "point") +
    scale_size_manual(
        NULL, values = 1.2, 
        guide = guide_legend(
            override.aes = list(linetype = "blank"))) +
    theme_bw() +
    theme(text = element_text(size = 10),
          legend.spacing.y = unit(0.1, 'cm'),
          plot.margin = unit(rep(0.1, 4), "cm"))

ggsave(file.path(fg_path, "figure1_study_area.png"), 
       width = 5, height = 3.8, dpi = 500, bg = "white")
```

## Results
### Figure 3

Figure 3 is the combination of predicted suitability maps.

```{r}
# Read layers
study_area <- read_sf(
    file.path(vct_path, "mainland_tanzania.geojson"))

suit_fine_scale <- read_stars(
    file.path(result_path, "landscape_utility_1km.tif")) %>% 
    split("band") %>% select("prediction")

suit_coarse_scale <- read_stars(
    file.path(result_path, "landscape_utility_10km_to_1km.tif")) %>% 
    split("band") %>% select("prediction")

suit_integrated <- read_stars(
    file.path(result_path, "landscape_utility_1km_integrated.tif")) %>% 
    split("band") %>% select("prediction")

suits <- c(suit_coarse_scale, suit_fine_scale, suit_integrated)
names(suits) <- c("Coarse", "Fine", "Integrated")

g1 <- ggplot() +
    geom_stars(data = suits %>% select("Coarse")) +
    scale_fill_viridis_c("Suitability\n(0-1)",
        option = "D", na.value = "transparent") +
    geom_sf(data = study_area, fill = "transparent", color = "black") +
    coord_sf() +
    theme_transparent() +
    theme(text = element_text(size = 12),
          legend.position = "bottom",
          legend.key.height = unit(0.3,"cm"),
          legend.text = element_text(size = 10),
          legend.margin = margin(0,0,0,0),
          legend.box.spacing = unit(0, "pt"),
          strip.text.x = element_text(size = 12, face = "bold"),
          plot.margin = unit(rep(-0.5, 4), "cm"))
g2 <- ggplot() +
    geom_stars(data = suits %>% select("Fine")) +
    scale_fill_viridis_c("Suitability\n(0-1)",
                         option = "D", na.value = "transparent") +
    geom_sf(data = study_area, fill = "transparent", color = "black") +
    coord_sf() +
    theme_transparent() +
    theme(text = element_text(size = 12),
          legend.position = "bottom",
          legend.key.height = unit(0.3,"cm"),
          legend.text = element_text(size = 10),
          legend.margin = margin(0,0,0,0),
          legend.box.spacing = unit(-0.1, "pt"),
          strip.text.x = element_text(size = 12, face = "bold"),
          plot.margin = unit(rep(-0.5, 4), "cm"))
g3 <- ggplot() +
    geom_stars(data = suits %>% select("Integrated")) +
    scale_fill_viridis_c("Suitability\n(0-1)",
                         option = "D", na.value = "transparent") +
    geom_sf(data = study_area, fill = "transparent", color = "black") +
    coord_sf() +
    theme_transparent() +
    theme(text = element_text(size = 12),
          legend.position = "bottom",
          legend.key.height = unit(0.3,"cm"),
          legend.text = element_text(size = 10),
          legend.margin = margin(rep(-1, 4)),
          legend.box.spacing = unit(-1, "pt"),
          legend.box.margin = margin(c(-20, 0, 0, 0)),
          strip.text.x = element_text(size = 12, face = "bold"),
          plot.margin = unit(c(-0.5, -0.5, 0, -0.5), "cm"))

ggarrange(ggarrange(g1 + theme(legend.position = "none"), 
          g2 + theme(legend.position = "none"), 
          ncol = 2, nrow = 1, labels = c("A", 'B'),
          font.label = list(size = 12, color = "black", face = "bold")),
          g3, ncol = 1, nrow = 2, labels = c("", 'C'),
          heights = c(1, 2),
          font.label = list(size = 12, color = "black", face = "bold"))

# ggarrange(g1, g2, g3, ncol = 1, nrow = 3, 
#           labels = c("A", 'B', 'C'),
#           font.label = list(size = 12, color = "black", face = "bold"),
#           common.legend = TRUE, legend = "bottom")

ggsave(file.path(fg_path, "figure3_suit_maps.png"), 
       width = 4, height = 6, dpi = 500, bg = "white")
```

### Statistics

```{r}
evals_5scale <- evals_train %>% filter(scale == 5) %>% 
    select(CBI, `AUC (ratio)`, TSS, AUC, `f-measure`)

ci <- function(v, conf_level = 0.95) {
    n <- length(v)
    se <- sd(v) / sqrt(n)
    qt(1 - ((1 - conf_level) / 2), n - 1) * se
}

evals_5scale <- evals_5scale %>%
    summarise_all(list(mean = mean, sd = sd, ci = ci))
```

```{r}
load(file.path(result_path, "runs_fs.rda"))
    
eval_fs <- do.call(rbind, lapply(runs, function(run) {
    tabularize_evaluation(run$eval_test)}))
rm(runs)

evals_fs <- eval_fs %>%
    summarise_all(list(mean = mean, sd = sd, ci = ci))
```

```{r}
load(file.path(result_path, "eval_fine_integrated.rda"))
eval_fine_integrated <- do.call(
    rbind, lapply(eval_fine_integrated, function(eval) {
    tabularize_evaluation(eval)}))

eval_fine_integrated <- eval_fine_integrated %>%
    summarise_all(list(mean = mean, sd = sd, ci = ci))
```

### Figure 4

```{r}
evals <- do.call(rbind, lapply(c(10, 5, 2.5), function(scale){
    do.call(rbind,lapply(c(0.1, 0.2, 0.3, 0.4, 0.5), function (ratio) {
        load(file.path(result_path, 
                       sprintf("tuning_cv_%s_%s.rda", scale, ratio)))
        
        tuning_cv %>% 
            select(sample_size, max_depth, ndim, scoring_metric,
                   cbi, auc_ratio, TSS, auc, `f-measure`) %>% 
            mutate(ratio = ratio)
    })) %>% mutate(scale = scale)
})) %>% rename(CBI = cbi, `AUC (ratio)` = auc_ratio, 
               AUC = auc, `F-measure` = `f-measure`)

# Check and compare the results
evals <- evals %>%
    pivot_longer(cols = c(`AUC (ratio)`, AUC, `F-measure`),
                 names_to = "metrics", values_to = "value") %>% 
    mutate(scale = factor(scale, levels = c(2.5, 5, 10)),
           ratio = factor(ratio, levels = c(0.1, 0.2, 0.3, 0.4, 0.5))) %>% 
    select(ratio, scale, metrics, value)

facet_names <- c("2.5 minutes", "5 minutes", "10 minutes")
names(facet_names) <- c("2.5", "5", "10")
ggplot(evals, 
       aes(x = ratio, y = value, fill = metrics, color = metrics)) +
    geom_boxplot(outlier.colour = "white", outlier.shape = 8,
                 outlier.size = 1, width = 0.5,
                 position = position_dodge(width = 0.9)) +
    stat_summary(
        fun = median,
        geom = 'line',
        size = 0.6, 
        aes(group = metrics, colour = metrics),
        position = position_dodge(width = 0.9)) +
    scale_fill_brewer("Evaluation metrics", palette = "Dark2",
                      labels = c("AUC",
                                  expression(paste(AUC[ratio])),
                                  "F-measure")) +
    scale_color_brewer("Evaluation metrics", palette = "Dark2",
                       labels = c("AUC",
                                  expression(paste(AUC[ratio])),
                                  "F-measure")) +
    stat_summary(
        fun = median,
        geom = 'point',
        color = "black",
        shape = "-",
        size = 5, 
        aes(group = metrics),
        position = position_dodge(width = 0.9),
        show.legend = FALSE) +
    facet_wrap(~scale, scales = "free_y", 
               labeller = labeller(scale = facet_names)) +
    ylab('') +
    xlab('Sampling ratio of pseudo occurrences') +
    theme_pubclean() +
    theme(text = element_text(size = 10, color = "black"),
          axis.text.x = element_text(angle = 0),
          axis.text = element_text(size = rel(0.8)),
          axis.title = element_text(size = rel(1)),
          legend.text = element_text(size = rel(1)),
          legend.position = "top",
          strip.text.x = element_text(size = 12, face = "bold")) +
    guides(fill = guide_legend(nrow = 1))

ggsave(file.path(fg_path, "figure4_optimal_pseudo_sample_size.png"), 
       width = 6, height = 3, dpi = 500)
```

### Figure 5

```{r}
library(tidyverse)
library(ggpubr)

tabularize_evaluation <- function(eval, type = "train") {
    if (type == "train") {
        tibble(
            CBI = eval$po_evaluation$boyce$cor,
            `AUC (ratio)` = eval$po_evaluation$roc_ratio$auc_ratio,
            TSS = eval$pb_evaluation$TSS$`Optimal TSS`,
            AUC = eval$pb_evaluation$roc$auc,
            Sensitivity = eval$pb_evaluation$sensitivity,
            `F-measure` = eval$pb_evaluation$`f-measure`)
    } else {
        tibble(
        Specificity = eval$pb_evaluation$specificity)
    }
}

evals_train <- do.call(rbind, lapply(c(10, 5, 2.5), function(scale){
    load(file.path(result_path, sprintf("runs_%s.rda", scale)))
    
    do.call(rbind, lapply(runs, function(run) {
        tabularize_evaluation(run$eval_test)
    })) %>% mutate(scale = scale)
}))

evals_sensitivity <- do.call(rbind, lapply(c(10, 5, 2.5), function(scale){
    load(file.path(result_path, sprintf("stt_real_same_%s.rda", scale)))
    
     do.call(rbind, lapply(sensitivity_real, function(each) {
        data.frame("Sensitivity" = each)
    })) %>% mutate(scale = scale)
}))

# Check and compare the results
evals_train <- evals_train %>% 
    pivot_longer(cols = c(`AUC (ratio)`, AUC, `F-measure`),
                 names_to = "metrics", values_to = "value") %>% 
    select(scale, metrics, value) %>% 
    mutate(scale = factor(
        scale, levels = c(2.5, 5, 10),
        labels = c("2.5 minutes", "5 minutes", "10 minutes")))

evals_sensitivity <- evals_sensitivity %>% 
    pivot_longer(cols = c(Sensitivity),
                 names_to = "metrics", values_to = "value") %>% 
    select(scale, metrics, value) %>% 
    mutate(scale = factor(
        scale, levels = c(2.5, 5, 10),
        labels = c("2.5 minutes", "5 minutes", "10 minutes")))

g_train <- ggplot(evals_train, 
       aes(x = metrics, y = value, fill = scale)) +
    geom_boxplot(outlier.colour = "black", outlier.alpha = 0,
                 outlier.size = 1) +
    scale_fill_brewer("Scale", palette = "Dark2") +
    scale_color_brewer("Scale", palette = "Dark2") +
    scale_x_discrete(limits = c("AUC", "AUC (ratio)", "F-measure"),
                     labels = c("AUC", expression("AUC"[ratio]), "F-measure")) +
    ylab('') +
    xlab('') +
    theme_classic() +
    theme(text = element_text(size = 10),
          axis.text.x = element_text(angle = 0),
          axis.text = element_text(size = rel(0.8)),
          axis.title = element_text(size = rel(1)),
          legend.text = element_text(size = rel(1)),
          legend.position = "top") +
    guides(fill = guide_legend(nrow = 1))

g_eval_stt <- ggplot(evals_sensitivity, 
                  aes(x = metrics, y = value, fill = scale)) +
    geom_boxplot(outlier.colour = "white", outlier.shape = 8,
                 outlier.size = 2) +
    scale_fill_brewer("Scale", palette = "Dark2") +
    ylab('') +
    xlab('') +
    theme_classic() +
    theme(text = element_text(size = 10),
          axis.text.x = element_text(angle = 0),
          axis.text = element_text(size = rel(0.8)),
          axis.title = element_text(size = rel(1)),
          legend.text = element_text(size = rel(1)),
          legend.position = "top") +
    guides(fill = guide_legend(nrow = 1))

figure <- ggarrange(g_train, g_eval_stt, 
                    labels = c("A", "B"),
                    font.label = list(size = 12, color = "black", 
                                      face = "bold"),
                    widths = c(2, 1),
                    common.legend = TRUE)

# Annotate the figure by adding a common labels
annotate_figure(figure,
                bottom = text_grob("Evaluation metrics", 
                                   color = "black", 
                                   size = 10, vjust = -1))

ggsave(file.path(fg_path, "figure5_eval_coarse_scales.png"), 
       width = 4, height = 3, dpi = 500, bg = "white")
```

### Figure 6

```{r}
library(tidytext)

# Fix a name lost bug here
nms <- c(paste0("BIO", c(1, 4, 7, 12, 15)), 
         "NDVI (dry season)", "NDVI (wet season)", "NDVI seasonality",
         "Road density", "River density", "Settlement density",
         "Surface roughness", "Patch density", "Patch richness density",
         "Shannon's diversity index", "Cropland edge density", 
         "Tree edge density", "Mean of Contiguity index (Savanna)", 
         "Savanna patch density", "Ratio of savanna", "Ratio of waterbodies")

abs_mean <- function(v) mean(abs(v))
var_imp <- do.call(rbind, lapply(c(10, 5, 2.5), function(scale){
    load(file.path(result_path, sprintf("runs_%s.rda", scale)))
    
    do.call(rbind, lapply(runs, function(run) {
        shap_train <- run$variable_analysis$SHAP$train %>% 
            summarise(across(all_of(names(.)), abs_mean))
        names(shap_train) <- nms
        shap_test <- run$variable_analysis$SHAP$test %>% 
            summarise(across(all_of(names(.)), abs_mean))
        names(shap_test) <- nms
        rbind(as_tibble(shap_train) %>% mutate(type = "train"),
        as_tibble(shap_test) %>% mutate(type = "test"))
    })) %>% mutate(scale = scale)
}))

## Reshape the table
var_imp <- var_imp %>% 
    pivot_longer(cols = rev(names(var_imp))[-c(1, 2)],
                 names_to = "variable", values_to = "value") %>% 
    mutate(scale = factor(scale, levels = c(2.5, 5, 10))) %>% 
    mutate(type = factor(type, levels = c("train", "test")))
var_imp <- var_imp %>% filter(type == "test") %>% 
    group_by(type, scale, variable) %>% 
    summarise(value = mean(value)) %>% ungroup()
var_imp_overall <- var_imp %>% group_by(variable) %>% 
    summarise(value = mean(value)) %>% 
    mutate(scale = "overall") %>% 
    mutate(value = value * 1000) %>% 
    arrange(value)
var_imp_overall <- var_imp_overall %>% 
    mutate(variable = factor(
        variable, levels = var_imp_overall$variable))
var_imp <- var_imp %>% group_by(scale, variable) %>% 
    summarise(value = mean(value)) %>% 
    mutate(value = value * 1000) %>% 
    mutate(scale = factor(
        scale, levels = c("2.5", "5", "10"),
        labels = c("2.5 minutes", 
                   "5 minutes", "10 minutes"))) %>% 
    mutate(variable = factor(
        variable, levels = var_imp_overall$variable))

ggplot(var_imp, aes(x = variable, y = value)) +
    geom_bar(data = var_imp_overall, aes(fill = "Overall"),
             stat = "identity", color = "transparent",
             position="identity") +
    geom_point(aes(color = scale), shape = "|", size = 4) +
    geom_point(aes(y = value + 0.01, color = scale), 
               shape = "|", size = 4) +
    geom_point(aes(y = value + 0.05, color = scale), 
               shape = "|", size = 4) +
    geom_point(aes(y = value + 0.09, color = scale), 
               shape = "|", size = 4) +
    scale_fill_manual(NULL, values = "black",
                      breaks = 'Overall') +
    scale_color_brewer("Scale", palette = "Dark2") +
    ylab('mean(|Shapley value|) * 1000') +
    xlab('Environmental variable') +
    coord_flip() +
    theme_pubclean() +
    theme(text = element_text(size = 10, color = "black"),
          axis.text.x = element_text(angle = 0),
          axis.text = element_text(size = rel(0.8), color = "black"),
          axis.title = element_text(size = rel(1)),
          legend.text = element_text(size = rel(1)),
          legend.position = "top", legend.justification = 'right',
          strip.text.x = element_text(size = 12, face = "bold")) +
    guides(fill = guide_legend(nrow = 1))

ggsave(file.path(fg_path, "figure6_var_imp_coarse.png"),
       width = 5, height = 4, dpi = 500)
```

### Figure 7

```{r}
library(ggpubr)

var_rsp <- do.call(rbind, lapply(c(10, 5, 2.5), function(scale){
    load(file.path(result_path, sprintf("runs_%s.rda", scale)))
    
    do.call(rbind, lapply(1:length(runs), function(n) {
        run <- runs[[n]]
        shaps <- run$shap_dependences$dependences_cont
        names(shaps) <- nms
        do.call(rbind, lapply(nms, function(nm) {
            shaps[[nm]] %>% mutate(variable = nm)
        })) %>% mutate(run = n)
    })) %>% mutate(scale = scale)
}))

# Calculate the density
for( var in c("Road density", "River density", "Settlement density")){
    for (sl in c(2.5, 5, 10)) {
        var_rsp <- var_rsp %>% 
                 mutate(x = ifelse(variable == var & scale == sl, 
                                   x / sl, x))}}

## Reformat the table
var_rsp <- var_rsp %>% 
    mutate(scale = factor(
        scale, levels = c(2.5, 5, 10),
        labels = c("2.5 minutes", "5 minutes", "10 minutes")))

## Plot (a template)
set.seed(123)
g_sd <- ggplot(var_rsp %>% 
           filter(variable == "Settlement density") %>% 
               group_by(scale) %>% 
               sample_n(size = 7000) %>%  
               ungroup()) +
    geom_point(aes(x = x, y = y, fill = scale), pch = 21,
               color = "white", size = 0.8, stroke = 0.2) +
    geom_smooth(aes(x = x, y = y, color = scale, fill = scale),
                se = TRUE, level = 0.95, size = 0.7,
                method = 'gam', formula = y ~ s(x, bs = "cs", k = 4)) +
    scale_color_brewer("Scale", palette = "Dark2") +
    scale_fill_brewer("Scale", palette = "Dark2") +
    geom_hline(yintercept = 0, linetype = "dashed") +
    scale_x_log10() +
    ylab('Shapley value') +
    xlab(expression(Settlement~density~(log[10]))) +
    theme_classic() +
    theme(text = element_text(size = 10, color = "black"),
          axis.text.x = element_text(angle = 0),
          axis.text = element_text(size = rel(0.8)),
          axis.title = element_text(size = rel(1)),
          legend.text = element_text(size = rel(1)),
          legend.position = "top") +
    guides(fill = guide_legend(nrow = 1))

set.seed(123)
g_bio4 <- ggplot(var_rsp %>% 
           filter(variable == "BIO4") %>% 
           group_by(scale) %>% 
               sample_n(size = 7000) %>% 
               ungroup()) +
    geom_point(aes(x = x, y = y, fill = scale), pch = 21,
               color = "white", size = 0.8, stroke = 0.2) +
    geom_smooth(aes(x = x, y = y, color = scale, fill = scale),
                se = TRUE, level = 0.95, method = 'gam', size = 0.7,
                formula = y ~ s(x, bs = "cs", k = 5)) +
    scale_color_brewer("Scale", palette = "Dark2") +
    scale_fill_brewer("Scale", palette = "Dark2") +
    geom_hline(yintercept = 0, linetype = "dashed") +
    ylab('Shapley value') +
    xlab("BIO4") +
    theme_classic() +
    theme(text = element_text(size = 10, color = "black"),
          axis.text.x = element_text(angle = 0),
          axis.text = element_text(size = rel(0.8)),
          axis.title = element_text(size = rel(1)),
          legend.text = element_text(size = rel(1)),
          legend.position = "top") +
    guides(fill = guide_legend(nrow = 1))

set.seed(123)
g_sdi <- ggplot(var_rsp %>% 
    filter(variable == "Shannon's diversity index") %>% 
    group_by(scale) %>% 
    sample_n(size = 7000) %>% 
    ungroup()) +
    geom_point(aes(x = x, y = y, fill = scale), pch = 21,
               color = "white", size = 0.8, stroke = 0.2) +
    geom_smooth(aes(x = x, y = y, color = scale, fill = scale),
                se = TRUE, level = 0.95, method = 'gam', size = 0.7,
                formula = y ~ s(x, bs = "cs", k = 4)) +
    scale_color_brewer("Scale", palette = "Dark2") +
    scale_fill_brewer("Scale", palette = "Dark2") +
    geom_hline(yintercept = 0, linetype = "dashed") +
    ylim(c(-0.08, 0.02)) +
    ylab('Shapley value') +
    xlab("Shannon's diversity index") +
    theme_classic() +
    theme(text = element_text(size = 10, color = "black"),
          axis.text.x = element_text(angle = 0),
          axis.text = element_text(size = rel(0.8)),
          axis.title = element_text(size = rel(1)),
          legend.text = element_text(size = rel(1)),
          legend.position = "top") +
    guides(fill = guide_legend(nrow = 1))

ggarrange(g_bio4, g_sd, g_sdi, ncol = 3,
          labels = c("A", "B", "C"),
          font.label = list(size = 12, color = "black", face = "bold"),
          common.legend = TRUE)

ggsave(file.path(fg_path, "figure7_variable_response.png"), 
       width = 6.5, height = 3, dpi = 500, bg = "white")
```

### Supplementary Figure S1-S3: Response curve of all variables at coarse scale

```{r}
# Types
type_a <- c(nms[1:8], "Cropland edge density", 
            "Patch density", "Ratio of waterbodies", 
            "Savanna patch density", "Tree edge density")
type_b <- c("River density", "Road density", "Settlement density",
            "Mean of Contiguity index (Savanna)")
type_c <- c("Shannon's diversity index", "Surface roughness",
            "Patch richness density", "Ratio of savanna")

# x value transition
group_orig <- c(nms[1:8], "Mean of Contiguity index (Savanna)",
                "Shannon's diversity index", "Ratio of savanna")
group_log <- c("Cropland edge density", 
               "Patch density", "Ratio of waterbodies", 
               "Savanna patch density", "Tree edge density",
               "River density", "Road density", 
               "Settlement density", "Surface roughness", 
               "Patch richness density")

## Plot lists
fg_orig_list <- lapply(group_orig, function(nm) {
    set.seed(123)
    
    if (nm == "Mean of Contiguity index (Savanna)") {
        data <- var_rsp %>% 
            filter(variable == nm) %>% 
            group_by(scale) %>% 
            sample_n(size = 7000) %>% 
            ungroup()
        ggplot(data) +
            geom_point(aes(x = x, y = y, fill = scale), pch = 21,
                       color = "white", size = 0.8, stroke = 0.2) +
            geom_smooth(data = data %>% filter(x > 0 & x < 0.95),
                        aes(x = x, y = y, color = scale, fill = scale),
                        se = TRUE, level = 0.95, method = 'gam', size = 0.7,
                        formula = y ~ s(x, bs = "cs", k = 4)) +
            scale_color_brewer("Scale", palette = "Dark2") +
            scale_fill_brewer("Scale", palette = "Dark2") +
            geom_hline(yintercept = 0, linetype = "dashed") +
            ylim(c(-0.08, 0.02)) +
            ylab('Shapley value') +
            xlab(expression(CONTIG[NM]~(savanna))) +
            theme_classic() +
            theme(text = element_text(size = 10, color = "black"),
                  axis.text.x = element_text(angle = 0),
                  axis.text = element_text(size = rel(0.8)),
                  axis.title = element_text(size = rel(1)),
                  legend.text = element_text(size = rel(1)),
                  legend.position = "top") +
            guides(fill = guide_legend(nrow = 1))
    } else if (nm == "Ratio of savanna") {
        data <- var_rsp %>% 
            filter(variable == nm) %>% 
            group_by(scale) %>% 
            sample_n(size = 7000) %>% 
            ungroup()
        ggplot(data) +
            geom_point(aes(x = x, y = y, fill = scale), pch = 21,
                       color = "white", size = 0.8, stroke = 0.2) +
            geom_smooth(data = data %>% filter(x < 0.99),
                        aes(x = x, y = y, color = scale, fill = scale),
                        se = TRUE, level = 0.95, size = 0.8,
                        method = 'gam', formula = y ~ s(x, bs = "cs", k = 5)) +
            scale_color_brewer("Scale", palette = "Dark2") +
            scale_fill_brewer("Scale", palette = "Dark2") +
            geom_hline(yintercept = 0, linetype = "dashed") +
            # scale_x_log10() +
            ylab('Shapley value') +
            xlab(nm) +
            theme_classic() +
            theme(text = element_text(size = 10, color = "black"),
                  axis.text.x = element_text(angle = 0),
                  axis.text = element_text(size = rel(0.8)),
                  axis.title = element_text(size = rel(1)),
                  legend.text = element_text(size = rel(1)),
                  legend.position = "top") +
            guides(fill = guide_legend(nrow = 1))
    } else {
        ggplot(var_rsp %>% 
                   filter(variable == nm) %>% 
                   group_by(scale) %>% 
                   sample_n(size = 7000) %>%  
                   ungroup()) +
            geom_point(aes(x = x, y = y, fill = scale), pch = 21,
                       color = "white", size = 0.8, stroke = 0.2) +
            geom_smooth(aes(x = x, y = y, color = scale, fill = scale),
                        se = TRUE, level = 0.95, size = 0.8,
                        method = 'gam', formula = y ~ s(x, bs = "cs", k = 6)) +
            scale_color_brewer("Scale", palette = "Dark2") +
            scale_fill_brewer("Scale", palette = "Dark2") +
            geom_hline(yintercept = 0, linetype = "dashed") +
            # scale_x_log10() +
            ylab('Shapley value') +
            xlab(nm) +
            theme_classic() +
            theme(text = element_text(size = 10, color = "black"),
                  axis.text.x = element_text(angle = 0),
                  axis.text = element_text(size = rel(0.8)),
                  axis.title = element_text(size = rel(1)),
                  legend.text = element_text(size = rel(1)),
                  legend.position = "top") +
            guides(fill = guide_legend(nrow = 1))
    }
})

names(fg_orig_list) <- group_orig

fg_log_list <- lapply(group_log, function(nm) {
    set.seed(123)
    
    nm_label <- parse(text = paste0(gsub(" ", "~", nm), "~''^'L'"))
    if (nm == "Patch richness density") {
        data <- var_rsp %>% 
            filter(variable == nm) %>% 
            group_by(scale) %>% 
            sample_n(size = 7000) %>%  
            ungroup()
        ggplot(data) +
            geom_point(aes(x = x, y = y, fill = scale), pch = 21,
                       color = "white", size = 0.8, stroke = 0.2) +
            geom_smooth(aes(x = x, y = y, color = scale, fill = scale),
                        se = TRUE, level = 0.95, size = 0.8) +
            scale_color_brewer("Scale", palette = "Dark2") +
            scale_fill_brewer("Scale", palette = "Dark2") +
            geom_hline(yintercept = 0, linetype = "dashed") +
            scale_x_log10() +
            ylab('Shapley value') +
            xlab(nm_label) +
            theme_classic() +
            theme(text = element_text(size = 10, color = "black"),
                  axis.text.x = element_text(angle = 0),
                  axis.text = element_text(size = rel(0.8)),
                  axis.title = element_text(size = rel(1)),
                  legend.text = element_text(size = rel(1)),
                  legend.position = "top") +
            guides(fill = guide_legend(nrow = 1))
    } else {
        ggplot(var_rsp %>% 
                   filter(variable == nm) %>% 
                   group_by(scale) %>% 
                   sample_n(size = 7000) %>%  
                   ungroup()) +
            geom_point(aes(x = x, y = y, fill = scale), pch = 21,
                       color = "white", size = 0.8, stroke = 0.2) +
            geom_smooth(aes(x = x, y = y, color = scale, fill = scale),
                        se = TRUE, level = 0.95, size = 0.8,
                        method = 'gam', formula = y ~ s(x, bs = "cs", k = 5)) +
            scale_color_brewer("Scale", palette = "Dark2") +
            scale_fill_brewer("Scale", palette = "Dark2") +
            geom_hline(yintercept = 0, linetype = "dashed") +
            scale_x_log10() +
            ylab('Shapley value') +
            xlab(nm_label) +
            theme_classic() +
            theme(text = element_text(size = 10, color = "black"),
                  axis.text.x = element_text(angle = 0),
                  axis.text = element_text(size = rel(0.8)),
                  axis.title = element_text(size = rel(1)),
                  legend.text = element_text(size = rel(1)),
                  legend.position = "top") +
            guides(fill = guide_legend(nrow = 1))
    }
})
names(fg_log_list) <- group_log

fg_list <- c(fg_orig_list, fg_log_list)

id_type_a <- which(names(fg_list) %in% type_a)
id_type_b <- which(names(fg_list) %in% type_b)
id_type_c <- which(names(fg_list) %in% type_c)
fg_type_a <- fg_list[id_type_a]
fg_type_b <- fg_list[id_type_b]
fg_type_c <- fg_list[id_type_c]

ggarrange(plotlist = fg_type_a, 
          ncol = 3, nrow = 5,
          common.legend = TRUE)

ggsave(file.path(fg_path, "sup_variable_response_a.png"), 
       width = 6.5, height = 8.5, dpi = 500, bg = "white")

ggarrange(plotlist = fg_type_b, 
          ncol = 2, nrow = 2,
          common.legend = TRUE)

ggsave(file.path(fg_path, "sup_variable_response_b.png"), 
       width = 4.3, height = 4, dpi = 500, bg = "white")

ggarrange(plotlist = fg_type_c, 
          ncol = 2, nrow = 2,
          common.legend = TRUE)

ggsave(file.path(fg_path, "sup_variable_response_c.png"), 
       width = 4.3, height = 4, dpi = 500, bg = "white")
```

### Figure 8

```{r}
# Define names
nms <- c("BIO4", "NDVI seasonality",
         "Distance to settlements", "Distance to waterbodies",
         "Vector ruggedness measure (3km)", 
         "Mean of Contiguity index (Savanna, 3km)", 
         "Cropland edge density (7km)", "Ratio of savanna (7km)",
         "Savanna patch density (3km)")

abs_mean <- function(v) mean(abs(v))
load(file.path(result_path, "runs_fs.rda"))
var_imp <- do.call(rbind, lapply(runs, function(run) {
    shap_test <- run$variable_analysis$SHAP$test %>% 
        summarise(across(all_of(names(.)), abs_mean))
    names(shap_test) <- nms
    shap_test
})); rm(runs)

## Reshape the table
var_imp <- var_imp %>% 
    pivot_longer(
        cols = names(var_imp),
        names_to = "variable", values_to = "value")
var_imp <- var_imp %>% 
    group_by(variable) %>% 
    summarise(value = mean(value)) %>% 
    ungroup() %>% arrange(value)
var_imp <- var_imp %>% 
    mutate(value = value * 1000) %>% 
    mutate(variable = factor(
        variable, levels = var_imp$variable))

ggplot(var_imp, aes(x = variable, y = value)) +
    geom_bar(fill = "black",
             stat = "identity", color = "transparent",
             position="identity") +
    ylab('mean(|Shapley value|) * 1000') +
    xlab('Environmental variable') +
    coord_flip() +
    theme_pubclean() +
    theme(text = element_text(size = 10, color = "black"),
          axis.text.x = element_text(angle = 0),
          axis.text = element_text(size = rel(0.8), color = "black"),
          axis.title = element_text(size = rel(1)),
          legend.text = element_text(size = rel(1)),
          legend.position = "top", legend.justification = 'right',
          strip.text.x = element_text(size = 12, face = "bold")) +
    guides(fill = guide_legend(nrow = 1))

ggsave(file.path(fg_path, "figure8_var_imp_fine.png"), 
       width = 5, height = 2, dpi = 500)
```

### Supplementary Figure S4: Response curve of all variables at fine scale

```{r}
# Define names
nms <- c("BIO4", "NDVI seasonality",
         "Distance to settlements", "Distance to waterbodies",
         "VRM (3km)", 
         "Mean of Contiguity index (Savanna, 3km)", 
         "ED of Cropland (7km)", "Ratio of savanna (7km)",
         "PD of Savanna (3km)")

load(file.path(result_path, "runs_fs.rda"))
var_rsp <- do.call(rbind, lapply(1:length(runs), function(n) {
        run <- runs[[n]]
        shaps <- run$shap_dependences$dependences_cont
        names(shaps) <- nms
        do.call(rbind, lapply(nms, function(nm) {
            shaps[[nm]] %>% mutate(variable = nm)
        })) %>% mutate(run = n)
    }))

fg_list <- lapply(nms, function(nm) {
    set.seed(123)
    
    if (nm %in% c("ED of Cropland (7km)",
                  "PD of Savanna (3km)",
                  "VRM (3km)")) {
        if (nm == "ED of Cropland (7km)"){
            nm_label <- expression(ED~of~Cropland~"(7km)"~""^"L")
        } else if (nm == "PD of Savanna (3km)") {
            nm_label <- expression(PD~of~Savanna~"(3km)"~""^"L")
        } else {
            nm_label <- expression(VRM~"(3km)"~""^"L")
        }
        ggplot(var_rsp %>% filter(variable == nm)) +
            geom_point(aes(x = x, y = y), pch = 21, fill = "black",
                       color = "white", size = 0.8, stroke = 0.2) +
            geom_smooth(aes(x = x, y = y),
                        se = TRUE, level = 0.95, size = 0.8,
                        method = 'gam', formula = y ~ s(x, bs = "cs", k = 4)) +
            scale_color_brewer("Scale", palette = "Dark2") +
            scale_fill_brewer("Scale", palette = "Dark2") +
            geom_hline(yintercept = 0, linetype = "dashed") +
            scale_x_log10() +
            ylab('Shapley value') +
            xlab(nm_label) +
            theme_classic() +
            theme(text = element_text(size = 9, color = "black"),
                  axis.text.x = element_text(angle = 0),
                  axis.text = element_text(size = rel(0.8)),
                  axis.title = element_text(size = rel(1)),
                  legend.text = element_text(size = rel(1)),
                  legend.position = "top") +
            guides(fill = guide_legend(nrow = 1))
    } else if (nm =="Mean of Contiguity index (Savanna, 3km)") {
            nm_label <- expression(CONTIG[NM]~of~Savanna~"(3km)")
            data <- var_rsp %>% filter(variable == nm)
            ggplot(data) +
                geom_point(aes(x = x, y = y), pch = 21, fill = "black",
                           color = "white", size = 0.8, stroke = 0.2) +
                geom_smooth(data = data %>% filter(x < 0.99 & x > 0.01),
                            aes(x = x, y = y),
                            se = TRUE, level = 0.95, size = 0.8,
                            method = 'gam', formula = y ~ s(x, bs = "cs", k = 6)) +
                geom_hline(yintercept = 0, linetype = "dashed") +
                ylab('Shapley value') +
                xlab(nm_label) +
                theme_classic() +
                theme(text = element_text(size = 9, color = "black"),
                      axis.text.x = element_text(angle = 0),
                      axis.text = element_text(size = rel(0.8)),
                      axis.title = element_text(size = rel(1)),
                      legend.text = element_text(size = rel(1)),
                      legend.position = "top") +
                guides(fill = guide_legend(nrow = 1))
            } else {
        ggplot(var_rsp %>% filter(variable == nm)) +
            geom_point(aes(x = x, y = y), pch = 21, fill = "black",
                       color = "white", size = 0.8, stroke = 0.2) +
            geom_smooth(aes(x = x, y = y),
                        se = TRUE, level = 0.95, size = 0.8,
                        method = 'gam', formula = y ~ s(x, bs = "cs", k = 6)) +
            geom_hline(yintercept = 0, linetype = "dashed") +
            ylab('Shapley value') +
            xlab(nm) +
            theme_classic() +
            theme(text = element_text(size = 9, color = "black"),
                  axis.text.x = element_text(angle = 0),
                  axis.text = element_text(size = rel(0.8)),
                  axis.title = element_text(size = rel(1)),
                  legend.text = element_text(size = rel(1)),
                  legend.position = "top") +
            guides(fill = guide_legend(nrow = 1))
    }
})
ggarrange(plotlist = fg_list, 
          ncol = 3, nrow = 3,
          common.legend = TRUE)

ggsave(file.path(fg_path, "sup_variable_response_fine.png"), 
       width = 6.5, height = 6.5, dpi = 500, bg = "white")
```

### Figure 9

Figure 9 is the landscape connectivity for long-distance migration.

```{r}
library(stars)
library(classInt)

# Define colors
col_habitats <- "#82ba9e"
col_blocks <- data.frame(
    id = 1:4,
    color = c("#fbf65d", "#fe0200", "#c99245", "#43642c"),
    name_= c("Farmland", "Settlement", "Farmland & settlement",
             "Mountain peak"))

# Read data
bry <- read_sf(file.path(vct_path, "mainland_tanzania.geojson"))
habitat_clusters <- read_stars(file.path(cnt_path, "habitat_clusters.tif"))
habitat_clusters[habitat_clusters == 0] <- NA
habitat_clusters[!is.na(habitat_clusters)] <- "Habitat cluster"

# Scenario 1
current_map_s1 <- read_stars(
    file.path(cnt_exp_path, "connectivity_scenario1.tif"))
blocks_s1 <- read_stars(
    file.path(cnt_path, "blocks_s1.tif")) %>% 
    st_crop(bry)
blocks_s1[blocks_s1 == 0] <- NA
names(blocks_s1) <- "blocks"
blocks_s1 <- blocks_s1 %>% 
    mutate(blocks = factor(
        blocks, levels = 1:4,
        labels = c("Farmland", "Settlement", 
                   "Farmland & settlement", "Mountain peaks"))) %>% 
    dplyr::select(blocks)

# Reformat the cuts to make the map looks pretty
values <- as.data.frame(current_map_s1) %>% na.omit() %>% pull(3)
values <- values[values >= quantile(values, 0.02) & 
                     values <= quantile(values, 0.98)]
breaks_s1 <- classIntervals(round(values, 3), 
                         n = 7, style = "equal")
labels <- sprintf("(%.3f,%.3f]", breaks_s1$brks, c(breaks_s1$brks[-1], 1))
current_map_s1 <- cut(current_map_s1, breaks = c(breaks_s1$brks, 1),
                      labels = labels)
# Plot
g_s1 <- ggplot() +
    geom_stars(data = current_map_s1, na.action = na.omit) +
    scale_fill_brewer(
        "Connectivity\nscenario 1\n(0-1)", direction = 1,
        palette = "BuPu",
        na.translate = FALSE) +
    guides(fill = guide_legend(nrow = 4, byrow = TRUE)) +
    ggnewscale::new_scale_fill() +
    geom_stars(data = habitat_clusters, 
               na.action = na.omit, show.legend = FALSE) +
    scale_fill_manual("", values = col_habitats,
                      na.translate = FALSE) +
    ggnewscale::new_scale_fill() +
    geom_stars(data = blocks_s1, show.legend = FALSE) +
    scale_fill_manual(
        "", values = col_blocks$color,
        na.translate = FALSE) +
    geom_sf(data = bry, color = "black", 
            fill = "transparent", linewidth = 0.5) +
    coord_sf() + theme_void() +
    theme(text = element_text(size = 10),
          legend.position = "bottom",
          legend.spacing.y = unit(0, 'cm'),
          plot.margin = unit(rep(0, 4), "cm"))

# Scenario 2
current_map_s2 <- read_stars(
    file.path(cnt_exp_path, "connectivity_scenario2.tif"))
blocks_s2 <- read_stars(
    file.path(cnt_path, "blocks_s2.tif")) %>% 
    st_crop(bry)
blocks_s2[blocks_s2 == 0] <- NA
names(blocks_s2) <- "blocks"
blocks_s2 <- blocks_s2 %>% 
    mutate(blocks = factor(
        blocks, levels = 1:4,
        labels = c("Farmland", "Settlement", 
                   "Farmland & settlement", "Mountain peaks"))) %>% 
    dplyr::select(blocks)

# Reformat the cuts to make the map looks pretty
values <- as.data.frame(current_map_s2) %>% na.omit() %>% pull(3)
values <- values[values >= quantile(values, 0.02) & 
                     values <= quantile(values, 0.98)]
breaks <- classIntervals(round(values, 3), 
                         n = 7, style = "equal")
labels <- sprintf("(%.3f,%.3f]", breaks$brks, c(breaks$brks[-1], 1))
current_map_s2 <- cut(current_map_s2, breaks = c(breaks$brks, 1),
                      labels = labels)
# Plot
g_s2 <- ggplot() +
    geom_stars(data = current_map_s2, na.action = na.omit) +
    scale_fill_brewer(
        "Connectivity\nscenario 2\n(0-1)", direction = 1,
        palette = "BuPu",
        na.translate = FALSE) +
    guides(fill = guide_legend(nrow = 4, byrow = TRUE)) +
    ggnewscale::new_scale_fill() +
    geom_stars(data = habitat_clusters, 
               na.action = na.omit, show.legend = FALSE) +
    scale_fill_manual("", values = col_habitats,
                      na.translate = FALSE) +
    ggnewscale::new_scale_fill() +
    geom_stars(data = blocks_s2, show.legend = FALSE) +
    scale_fill_manual(
        "", values = col_blocks$color,
        na.translate = FALSE) +
    geom_sf(data = bry, color = "black", 
            fill = "transparent", linewidth = 0.5) +
    coord_sf() + theme_void() +
    theme(text = element_text(size = 10),
          legend.position = "bottom",
          legend.spacing.y = unit(0, 'cm'),
          plot.margin = unit(rep(0, 4), "cm"))

# Legend
current_map_s1 <- read_stars(
    file.path(cnt_exp_path, "connectivity_scenario1.tif"))
legend_common <- ggplot() +
    geom_stars(data = current_map_s1, 
               na.action = na.omit) +
    scale_fill_distiller("Connectivity",
        palette = "BuPu", breaks = c(0, 0.5, 1), 
        labels = c("low", "med", "high"), direction = 1) +
    ggnewscale::new_scale_fill() +
    geom_stars(data = habitat_clusters, 
               na.action = na.omit) +
    scale_fill_manual("", values = col_habitats,
                      na.translate = FALSE) +
    ggnewscale::new_scale_fill() +
    geom_stars(data = blocks_s2) +
    scale_fill_manual(
        "", values = col_blocks$color,
        na.translate = FALSE) +
    theme(text = element_text(size = 10),
          legend.position = "bottom",
          legend.spacing.y = unit(0.1, 'cm'),
          legend.box = "vertical",
          plot.margin = unit(rep(0, 4), "cm"),
          legend.margin = margin(0,0,0,0),
          legend.box.margin = margin(0,0,0,0))
legend_common <- get_legend(legend_common)

# Combining plots
plots <- ggarrange(g_s1 + theme(legend.position = "none"), 
                   g_s2 + theme(legend.position = "none"), 
                   labels = c("A", "B"),
                   font.label = list(size = 12),
                   nrow = 1)
ggarrange(plots, ggplot() + theme_void(),
          legend_common, ggplot() + theme_void(), nrow = 4, 
          heights = c(0.9, -0.01, 0.1, 0.09))

ggsave(file.path(fg_path, "figure9_long_distance_connect.png"), 
       bg = "white",
       width = 6.5, height = 4.2, dpi = 500)
```

### Figure 10

Figure 10 is the main figure for the detected corridors and example path across corridors.

```{r}
# Main figure
bry <- read_sf(file.path(vct_path, "mainland_tanzania.geojson"))
habitat_clusters <- read_sf(file.path(data_path, "observations",
                            "habitat_clusters_aggre.geojson"))
corridors <- read_sf(file.path(cnt_exp_path, "corridors_s1_simple.geojson"))
corridors <- st_join(corridors, random_paths)

ggplot() +
    geom_sf(data = habitat_clusters, aes(fill = "Habitat clusters"),
            color = "transparent") +
    geom_sf_text(data = habitat_clusters,
                 aes(label = cluster), color = "black", 
                 show_guide = FALSE) +
    geom_sf(data = corridors, aes(fill = "Corridors"), 
            color = "transparent") +
    scale_fill_manual(NULL, values = c(col_habitats, "gray50"),
                      breaks = c("Habitat clusters", 'Corridors'),
                      guide = guide_legend(
                          override.aes = list(linetype = "blank", 
                                              shape = NA))) +
    geom_sf(data = random_paths, aes(color = id), linewidth = 0.4) +
    scale_color_brewer("", palette = "Set1", direction = -1,
                       guide = "none") +
    geom_sf_text(data = random_paths,
                 aes(label = id), color = "black", 
                 show_guide = FALSE) +
    geom_sf(data = bry, fill = "transparent", color = "black",
            linewidth = 0.2) +
    theme_void() +
    theme(text = element_text(size = 12),
          legend.spacing.y = unit(0, 'cm'),
          plot.margin = unit(rep(0, 4), "cm"),
          legend.position = "bottom")

ggsave(file.path(fg_path, "figure10_corridors.png"), 
       bg = "white",
       width = 3, height = 3, dpi = 500)
```


### Figure 11

Figure 11 is the environmental variable values across the random paths shown in figure 10.

```{r}
load(file.path(cnt_exp_path, "random_path_statistics.rda"))
random_paths <- read_sf(
    file.path(cnt_exp_path, "random_path_within_corridor.shp")) %>% 
    mutate(id = LETTERS[1:6])

norm_data <- function(s){
    (s - min(s)) / (max(s) - min(s))
}

fg_list <- lapply(random_path_statistics, function(each){
    x_nm <- ifelse(names(each)[1] == "lat", "Latitude", "Longitude")
    vals <- each
    if ( x_nm == "Latitude") {
        ele_dif <- max(vals[vals$variable == "Elevation", "value"]) - 
            min(vals[vals$variable == "Elevation", "value"])
        vals[vals$variable == "Elevation", "value"] <- 
            norm_data(vals[vals$variable == "Elevation", "value"])
        ggplot() +
            geom_area(data = vals %>% filter(variable == "Elevation"),
                      aes(x = lat, y = value, fill = "Elevation"), 
                      color = "transparent",
                      show.legend = TRUE) +
            scale_fill_manual(NULL, values = "grey",
                              breaks = 'Elevation',
                              guide = guide_legend(
                                  override.aes = list(linetype = "blank", 
                                                      shape = NA))) +
            geom_line(data = vals %>% filter(variable != "Elevation"), 
                      aes(x = lat, y = value, color = variable),
                      linewidth = 0.6) +
            scale_color_brewer("", palette = "Dark2",
                               guide = guide_legend(
                                   override.aes = list(fill = NA))) +
            scale_y_continuous(
                "Cropland coverage", 
                sec.axis = sec_axis(~ . * 1, name = "NDVI")
            ) + xlab(x_nm) +
            annotate(geom = "text", x = min(vals[, 1]) - 0.05, y = 0.5, 
                     label = sprintf("Elevation difference: %.1f m", ele_dif), 
                     color = "red", size = 2.5, angle = 90) +
            theme_classic() +
            theme(legend.position = "top")+
            theme(text = element_text(size = 10),
                  legend.position = "bottom")
    } else {
        ele_dif <- max(vals[vals$variable == "Elevation", "value"]) - 
            min(vals[vals$variable == "Elevation", "value"])
        vals[vals$variable == "Elevation", "value"] <- 
            norm_data(vals[vals$variable == "Elevation", "value"])
            
        ggplot() +
            geom_area(data = vals %>% filter(variable == "Elevation"),
                      aes(x = lon, y = value, fill = "Elevation"), 
                      color = "transparent",
                      show.legend = TRUE) +
            scale_fill_manual(NULL, values = "grey",
                              breaks = 'Elevation',
                              guide = guide_legend(
                                  override.aes = list(linetype = "blank", 
                                                      shape = NA))) +
            geom_line(data = vals %>% filter(variable != "Elevation"), 
                      aes(x = lon, y = value, color = variable),
                      linewidth = 0.6) +
            scale_color_brewer("", palette = "Dark2",
                               guide = guide_legend(
                                   override.aes = list(fill = NA))) +
            scale_y_continuous(
                "Cropland coverage", 
                sec.axis = sec_axis(~ . * 1, name = "NDVI")) + xlab(x_nm) +
            annotate(geom = "text", x = min(vals[, 1]) - 0.03, y = 0.5, 
                     label = sprintf("Elevation difference: %.1f m", ele_dif), 
                     color = "red", size = 2.5, angle = 90) +
            theme_classic() +
            theme(legend.position = "top") +
            theme(text = element_text(size = 10),
                  legend.position = "bottom")
    }
})

ggarrange(plotlist = fg_list, nrow = 3, ncol = 2,
          labels = LETTERS[1:6], font.label = list(size = 12),
          common.legend = TRUE, legend = "top")

ggsave(file.path(fg_path, "figure11_path_analysis.png"), 
       bg = "white",
       width = 6.5, height = 6.2, dpi = 500)
```

### Figure 12

Figure 12 shows the simulated landscape connectivity for short-distance move.

#### Figure 12-1 

The figure for habitat cluster 1.

```{r}
library(classInt)
cls_path <- file.path(cnt_exp_path, "cluster_1")

# Borders
pas <- read_sf(file.path(cnt_path, "pas.geojson"))
pas_c1 <- pas %>% filter(cluster == 1)
bry <- st_union(pas_c1) %>% st_convex_hull() %>% 
    st_as_sf() %>% rename(geometry = x)
hotspots_c1 <- read_sf(file.path(cls_path, "hotspots_conservation.shp"))

# Scenario 1
current_map_s1 <- read_stars(
    file.path(cls_path, "c1_connectivity_scenario1.tif"))
blocks_s1 <- read_stars(
    file.path(cnt_path, "blocks_spath_s1.tif")) %>% 
    st_crop(bry)
blocks_s1[blocks_s1 == 0] <- NA
names(blocks_s1) <- "blocks"
blocks_s1 <- blocks_s1 %>% 
    mutate(blocks = factor(
        blocks, levels = 1:4,
        labels = c("Farmland", "Settlement", 
                   "Farmland & settlement", "Mountain peaks"))) %>% 
    dplyr::select(blocks)

# Reformat the cuts to make the map looks pretty
values <- as.data.frame(current_map_s1) %>% na.omit() %>% pull(3)
values <- values[values >= quantile(values, 0.01) & 
                     values <= quantile(values, 0.99)]
breaks_s1 <- classIntervals(round(values, 3), 
                         n = 7, style = "equal")
labels <- sprintf("(%.2f,%.2f]", breaks_s1$brks, c(breaks_s1$brks[-1], 1))
current_map_s1 <- cut(current_map_s1, breaks = c(breaks_s1$brks, 1),
                      labels = labels)
# Plot
g_s1 <- ggplot() +
    geom_stars(data = current_map_s1, na.action = na.omit) +
    scale_fill_brewer(
        "Connectivity\nscenario 1\n(0-1)", direction = 1,
        palette = "BuPu",
        na.translate = FALSE) +
    guides(fill = guide_legend(nrow = 4, byrow = TRUE)) +
    ggnewscale::new_scale_fill() +
    geom_stars(data = blocks_s1, show.legend = FALSE) +
    scale_fill_manual(
        "", values = col_blocks$color,
        na.translate = FALSE) +
    geom_sf(data = pas_c1, color = "gray40", 
            fill = "transparent") +
    geom_sf(data = bry, color = "black", 
            fill = "transparent") +
    coord_sf() + theme_void() +
    theme(text = element_text(size = 10),
          legend.position = "bottom",
          legend.spacing.y = unit(0, 'cm'),
          plot.margin = unit(rep(0, 4), "cm"))

# Scenario 2
current_map_s2 <- read_stars(
    file.path(cls_path, "c1_connectivity_scenario2.tif"))
blocks_s2 <- read_stars(
    file.path(cnt_path, "blocks_spath_s2.tif")) %>% 
    st_crop(bry)
blocks_s2[blocks_s2 == 0] <- NA
names(blocks_s2) <- "blocks"
blocks_s2 <- blocks_s2 %>% 
    mutate(blocks = factor(
        blocks, levels = 1:4,
        labels = c("Farmland", "Settlement", 
                   "Farmland & settlement", "Mountain peaks"))) %>% 
    dplyr::select(blocks)

# Reformat the cuts to make the map looks pretty
values <- as.data.frame(current_map_s2) %>% na.omit() %>% pull(3)
values <- values[values >= quantile(values, 0.01) & 
                     values <= quantile(values, 0.99)]
breaks <- classIntervals(round(values, 3), 
                         n = 7, style = "equal")
labels <- sprintf("(%.2f,%.2f]", breaks$brks, c(breaks$brks[-1], 1))
current_map_s2 <- cut(current_map_s2, breaks = c(breaks$brks, 1),
                      labels = labels)
# Plot
g_s2 <- ggplot() +
    geom_stars(data = current_map_s2, na.action = na.omit) +
    scale_fill_brewer(
        "Connectivity\nscenario 1\n(0-1)", direction = 1,
        palette = "BuPu",
        na.translate = FALSE) +
    guides(fill = guide_legend(nrow = 4, byrow = TRUE)) +
    ggnewscale::new_scale_fill() +
    geom_stars(data = blocks_s2, show.legend = FALSE) +
    scale_fill_manual(
        "", values = col_blocks$color,
        na.translate = FALSE) +
    geom_sf(data = pas_c1, color = "gray40", 
            fill = "transparent") +
    geom_sf(data = hotspots_c1, color = "blue", 
            fill = "transparent", linewidth = 0.8) +
    geom_sf(data = bry, color = "black", 
            fill = "transparent") +
    coord_sf() + theme_void() +
    theme(text = element_text(size = 10),
          legend.position = "bottom",
          legend.spacing.y = unit(0, 'cm'),
          plot.margin = unit(rep(0, 4), "cm"))

plot_c1 <- ggarrange(g_s1 + theme(legend.position = "none"), 
                     g_s2 + theme(legend.position = "none"), 
                     labels = c("1.A", "1.B"),
                     font.label = list(size = 12),
                     nrow = 2)
```

#### Figure 12-2

The figure for habitat Cluster 2.

```{r}
cls_path <- file.path(cnt_exp_path, "cluster_2")

# Borders
pas <- read_sf(file.path(cnt_path, "pas.geojson"))
pas_c2 <- pas %>% filter(cluster == 2)
bry <- st_union(pas_c2) %>% st_convex_hull() %>% 
    st_as_sf() %>% rename(geometry = x)
hotspots_c2 <- read_sf(file.path(cls_path, "hotspots_conservation_c2.shp"))

# Scenario 1
current_map_s1 <- read_stars(
    file.path(cls_path, "c2_connectivity_scenario1.tif"))
blocks_s1 <- read_stars(
    file.path(cnt_path, "blocks_spath_s1.tif")) %>% 
    st_crop(bry)
blocks_s1[blocks_s1 == 0] <- NA
names(blocks_s1) <- "blocks"
blocks_s1 <- blocks_s1 %>% 
    mutate(blocks = factor(
        blocks, levels = 1:4,
        labels = c("Farmland", "Settlement", 
                   "Farmland & settlement", "Mountain peaks"))) %>% 
    dplyr::select(blocks)

# Reformat the cuts to make the map looks pretty
values <- as.data.frame(current_map_s1) %>% na.omit() %>% pull(3)
values <- values[values >= quantile(values, 0.01) & 
                     values <= quantile(values, 0.99)]
breaks_s1 <- classIntervals(round(values, 3), 
                         n = 7, style = "equal")
labels <- sprintf("(%.2f,%.2f]", breaks_s1$brks, c(breaks_s1$brks[-1], 1))
current_map_s1 <- cut(current_map_s1, breaks = c(breaks_s1$brks, 1),
                      labels = labels)
# Plot
g_s1 <- ggplot() +
    geom_stars(data = current_map_s1, na.action = na.omit) +
    scale_fill_brewer(
        "Connectivity\nscenario 1\n(0-1)", direction = 1,
        palette = "BuPu",
        na.translate = FALSE) +
    guides(fill = guide_legend(nrow = 4, byrow = TRUE)) +
    ggnewscale::new_scale_fill() +
    geom_stars(data = blocks_s1, show.legend = FALSE) +
    scale_fill_manual(
        "", values = col_blocks$color,
        na.translate = FALSE) +
    geom_sf(data = pas_c2, color = "gray40", 
            fill = "transparent") +
    geom_sf(data = bry, color = "black", 
            fill = "transparent") +
    coord_sf() + theme_void() +
    theme(text = element_text(size = 10),
          legend.position = "bottom",
          legend.spacing.y = unit(0, 'cm'),
          plot.margin = unit(rep(0, 4), "cm"))

# Scenario 2
current_map_s2 <- read_stars(
    file.path(cls_path, "c2_connectivity_scenario2.tif"))
blocks_s2 <- read_stars(
    file.path(cnt_path, "blocks_spath_s2.tif")) %>% 
    st_crop(bry)
blocks_s2[blocks_s2 == 0] <- NA
names(blocks_s2) <- "blocks"
blocks_s2 <- blocks_s2 %>% 
    mutate(blocks = factor(
        blocks, levels = 1:4,
        labels = c("Farmland", "Settlement", 
                   "Farmland & settlement", "Mountain peaks"))) %>% 
    dplyr::select(blocks)

# Reformat the cuts to make the map looks pretty
values <- as.data.frame(current_map_s2) %>% na.omit() %>% pull(3)
values <- values[values >= quantile(values, 0.01) & 
                     values <= quantile(values, 0.99)]
breaks <- classIntervals(round(values, 3), 
                         n = 7, style = "equal")
labels <- sprintf("(%.2f,%.2f]", breaks$brks, c(breaks$brks[-1], 1))
current_map_s2 <- cut(current_map_s2, breaks = c(breaks$brks, 1),
                      labels = labels)
# Plot
g_s2 <- ggplot() +
    geom_stars(data = current_map_s2, na.action = na.omit) +
    scale_fill_brewer(
        "Connectivity\nscenario 1\n(0-1)", direction = 1,
        palette = "BuPu",
        na.translate = FALSE) +
    guides(fill = guide_legend(nrow = 4, byrow = TRUE)) +
    ggnewscale::new_scale_fill() +
    geom_stars(data = blocks_s2, show.legend = FALSE) +
    scale_fill_manual(
        "", values = col_blocks$color,
        na.translate = FALSE) +
    geom_sf(data = pas_c2, color = "gray40", 
            fill = "transparent") +
    geom_sf(data = hotspots_c2, color = "blue", 
            fill = "transparent", linewidth = 0.8) +
    geom_sf(data = bry, color = "black", 
            fill = "transparent") +
    coord_sf() + theme_void() +
    theme(text = element_text(size = 10),
          legend.position = "bottom",
          legend.spacing.y = unit(0, 'cm'),
          plot.margin = unit(rep(0, 4), "cm"))

plot_c2 <- ggarrange(g_s1 + theme(legend.position = "none"), 
                     g_s2 + theme(legend.position = "none"), 
                     labels = c("2.A", "2.B"),
                     font.label = list(size = 12),
                     nrow = 2)
```

#### Figure 12-3

The figure for habitat Cluster 3.

```{r}
cls_path <- file.path(cnt_exp_path, "cluster_3")

# Borders
pas <- read_sf(file.path(cnt_path, "pas.geojson"))
pas_c3 <- pas %>% filter(cluster == 3)
bry <- st_union(pas_c3) %>% st_convex_hull() %>% 
    st_as_sf() %>% rename(geometry = x)
hotspots_c3 <- read_sf(file.path(cls_path, "hotspots_conservation_c3.shp"))

# Scenario 1
current_map_s1 <- read_stars(
    file.path(cls_path, "c3_connectivity_scenario1.tif"))
blocks_s1 <- read_stars(
    file.path(cnt_path, "blocks_spath_s1.tif")) %>% 
    st_crop(bry)
blocks_s1[blocks_s1 == 0] <- NA
names(blocks_s1) <- "blocks"
blocks_s1 <- blocks_s1 %>% 
    mutate(blocks = factor(
        blocks, levels = 1:4,
        labels = c("Farmland", "Settlement", 
                   "Farmland & settlement", "Mountain peaks"))) %>% 
    dplyr::select(blocks)

# Reformat the cuts to make the map looks pretty
values <- as.data.frame(current_map_s1) %>% na.omit() %>% pull(3)
values <- values[values >= quantile(values, 0.01) & 
                     values <= quantile(values, 0.99)]
breaks_s1 <- classIntervals(round(values, 3), 
                         n = 7, style = "equal")
labels <- sprintf("(%.2f,%.2f]", breaks_s1$brks, c(breaks_s1$brks[-1], 1))
current_map_s1 <- cut(current_map_s1, breaks = c(breaks_s1$brks, 1),
                      labels = labels)
# Plot
g_s1 <- ggplot() +
    geom_stars(data = current_map_s1, na.action = na.omit) +
    scale_fill_brewer(
        "Connectivity\nscenario 1\n(0-1)", direction = 1,
        palette = "BuPu",
        na.translate = FALSE) +
    guides(fill = guide_legend(nrow = 4, byrow = TRUE)) +
    ggnewscale::new_scale_fill() +
    geom_stars(data = blocks_s1, show.legend = FALSE) +
    scale_fill_manual(
        "", values = col_blocks$color,
        na.translate = FALSE) +
    geom_sf(data = pas_c3, color = "gray40", 
            fill = "transparent") +
    geom_sf(data = bry, color = "black", 
            fill = "transparent") +
    coord_sf() + theme_void() +
    theme(text = element_text(size = 10),
          legend.position = "bottom",
          legend.spacing.y = unit(0, 'cm'),
          plot.margin = unit(rep(0, 4), "cm"))

# Scenario 2
current_map_s2 <- read_stars(
    file.path(cls_path, "c3_connectivity_scenario2.tif"))
blocks_s2 <- read_stars(
    file.path(cnt_path, "blocks_spath_s2.tif")) %>% 
    st_crop(bry)
blocks_s2[blocks_s2 == 0] <- NA
names(blocks_s2) <- "blocks"
blocks_s2 <- blocks_s2 %>% 
    mutate(blocks = factor(
        blocks, levels = 1:4,
        labels = c("Farmland", "Settlement", 
                   "Farmland & settlement", "Mountain peaks"))) %>% 
    dplyr::select(blocks)

# Reformat the cuts to make the map looks pretty
values <- as.data.frame(current_map_s2) %>% na.omit() %>% pull(3)
values <- values[values >= quantile(values, 0.01) & 
                     values <= quantile(values, 0.99)]
breaks <- classIntervals(round(values, 3), 
                         n = 7, style = "equal")
labels <- sprintf("(%.2f,%.2f]", breaks$brks, c(breaks$brks[-1], 1))
current_map_s2 <- cut(current_map_s2, breaks = c(breaks$brks, 1),
                      labels = labels)
# Plot
g_s2 <- ggplot() +
    geom_stars(data = current_map_s2, na.action = na.omit) +
    scale_fill_brewer(
        "Connectivity\nscenario 1\n(0-1)", direction = 1,
        palette = "BuPu",
        na.translate = FALSE) +
    guides(fill = guide_legend(nrow = 4, byrow = TRUE)) +
    ggnewscale::new_scale_fill() +
    geom_stars(data = blocks_s2, show.legend = FALSE) +
    scale_fill_manual(
        "", values = col_blocks$color,
        na.translate = FALSE) +
    geom_sf(data = pas_c3, color = "gray40", 
            fill = "transparent") +
    geom_sf(data = hotspots_c3, color = "blue", 
            fill = "transparent", linewidth = 0.8) +
    geom_sf(data = bry, color = "black", 
            fill = "transparent") +
    coord_sf() + theme_void() +
    theme(text = element_text(size = 10),
          legend.position = "bottom",
          legend.spacing.y = unit(0, 'cm'),
          plot.margin = unit(rep(0, 4), "cm"))

plot_c3 <- ggarrange(g_s1 + theme(legend.position = "none"), 
                     g_s2 + theme(legend.position = "none"), 
                     labels = c("3.A", "3.B"),
                     font.label = list(size = 12),
                     nrow = 2)
```

#### Plot together

```{r}
# Make the common legend
current_map_s1 <- read_stars(
    file.path(cls_path, "c3_connectivity_scenario1.tif"))
legend_common <- ggplot() +
    geom_stars(data = current_map_s1, 
               na.action = na.omit) +
    scale_fill_distiller("Connectivity",
        palette = "BuPu", breaks = c(0, 0.5, 1), 
        labels = c("low", "med", "high"), direction = 1) +
    ggnewscale::new_scale_fill() +
    geom_stars(data = blocks_s2) +
    scale_fill_manual(
        "", values = col_blocks$color,
        na.translate = FALSE) +
    geom_sf(data = pas_c3, aes(color = "Protected areas"), 
            fill = "transparent") +
    geom_sf(data = bry, aes(color = "Critical regions"), 
            fill = "transparent") +
    scale_color_manual(NULL, values = c("gray40", "blue"),
                      breaks = c('Protected areas', "Critical regions")) +
    theme(text = element_text(size = 10),
          legend.position = "bottom",
          legend.spacing.y = unit(0.1, 'cm'),
          legend.box = "vertical",
          plot.margin = unit(rep(0, 4), "cm"),
          legend.margin = margin(0,0,0,0),
          legend.box.margin = margin(0,0,0,0))
legend_common <- get_legend(legend_common)

ggarrange(ggarrange(plot_c1, plot_c2, 
                    plot_c3, ncol = 3), 
          ggplot() + theme_void(),
          legend_common, ggplot() + theme_void(),
          nrow = 4, heights = c(3, -0.3, 1, -0.2))

ggsave(file.path(fg_path, "figure12_clusters_condition.png"), 
       bg = "white",
       width = 6.5, height = 6, dpi = 500)
```

